---
title: "Data 607 - Final Project - DB Load"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Import Libraries
```{r}
library(tidyverse)
library(DBI)
library(odbc)
```


## Load Data
Movie rating data is available at https://grouplens.org/datasets/movielens/25m/ which is produced by GroupLens. From their website,
"GroupLens is a research lab in the Department of Computer Science and Engineering at the University of Minnesota, Twin Cities specializing in recommender systems, online communities, mobile and ubiquitous technologies, digital libraries, and local geographic information systems."

```{r}
movie_df <- readr::read_csv("movies.csv", col_types = "icc")
link_data <- readr::read_csv("links.csv", col_select = -tmdbId, col_types = "ic")
rating_df <- readr::read_csv("ratings.csv", col_types = "iid?")
```


## Data Munging

```{r}
movie_df <- movie_df|> dplyr::rename(movieid = movieId)
link_data <- link_data|> dplyr::rename(c(movieid = movieId, imdbid = imdbId))
rating_df <- rating_df|> dplyr::rename(c(movieid = movieId, userid = userId))

movie_df <- movie_df|> dplyr::left_join(link_data, by="movieid")
movie_df <- movie_df|> dplyr::mutate(imdbid = paste0("tt",movie_df$imdbid)) |>
            dplyr::mutate(year = as.integer(stringr::str_sub(movie_df$title,-5,-2))) |>
            dplyr::mutate(title = str_sub(movie_df$title, 1, -8)) |>
            dplyr::filter(year >= 2010)

#remove ratings below 3 because AWS Personalize can't process negative feedback
rating_df <- rating_df |> dplyr::filter(rating>=3)

rating_count_bymovie <- movie_df |> left_join(rating_df, by='movieid' ) |> group_by(movieid) |> tally(!is.na(userid), name="rating_count")

#subset movie datafframe for the 2500 most reviewed movie titles 
movie_df <- movie_df |> inner_join(rating_count_bymovie, by="movieid") |>
          dplyr::slice_max(order_by = rating_count, n = 2500, with_ties = FALSE)

movie_genre_df <- dplyr::select(movie_df, movieid, genres)
movie_genre_df <- movie_genre_df |> dplyr::rename(genre = genres) |>
            tidyr::separate_rows(genre, sep="\\|") |>
            dplyr::filter(genre != "(no genres listed)")

genre_df <- dplyr::distinct(movie_genre_df, genre) |>
            dplyr::mutate(genreid = row_number()) |>
            dplyr::relocate(genreid, .before = "genre")

movie_genre_df <- movie_genre_df |> dplyr::inner_join(genre_df, by="genre") |>
                  dplyr::select(-genre)

movie_df <- dplyr::select(movie_df, -c(genres, rating_count))
```

```{r}
rating_df <- rating_df |> dplyr::semi_join(movie_df, by="movieid")
rating_df$timestamp = as.POSIXct(rating_df$timestamp, origin = "1970-01-01")
```

## Summerize Interaction Data
Showing the thousands of movies that were excluded for having low user rating counts.

```{r}
rating_count_bymovie |> summarise(median = median(rating_count), average= mean(rating_count), count = n())

rating_count_low <- rating_count_bymovie |> dplyr::filter(rating_count <10)
ggplot(data = rating_count_low, aes(x=rating_count)) + geom_histogram()
```


## Database Connection

You must have RMySQL package installed.

```{r}
#Establish connection to DB
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "moviedb", 
                 host = D607_FINAL_ENV$AWS_RDS_HOST, 
                 port = 3306,
                 user = D607_FINAL_ENV$AWS_RDS_USER_NAME,
                 password = D607_FINAL_ENV$AWS_RDS_PW)
```

## Load DB Data

```{r}
#load movies
dbWriteTable(con, "MOVIE", movie_df, overwrite = FALSE, append = TRUE, row.names = FALSE) #load into DB
test_movie <- odbc::dbGetQuery(con, "SELECT * FROM moviedb.MOVIE LIMIT 6;") #test insert
print(test_movie)

#load user ratings
dbWriteTable(con, "RATING", rating_df, overwrite = FALSE, append = TRUE, row.names = FALSE) #load into DB
test_rating <- odbc::dbGetQuery(con, "SELECT * FROM moviedb.RATING LIMIT 6;") #test insert
print(test_rating)

#load genre
dbWriteTable(con, "GENRE", genre_df, overwrite = FALSE, append = TRUE, row.names = FALSE) #load into DB
test_genre <- odbc::dbGetQuery(con, "SELECT * FROM moviedb.GENRE LIMIT 6;") #test insert
print(test_genre)

#load movie genre join table
dbWriteTable(con, "MOVIE_GENRE", movie_genre_df, overwrite = FALSE, append = TRUE, row.names = FALSE) #load into DB
test_movie_genre <- odbc::dbGetQuery(con, "SELECT * FROM moviedb.MOVIE_GENRE LIMIT 6;") #test insert
print(test_movie_genre)
```

